---
title: "Modifying forecast's tsCV"
author: "John Grando"
date: "October 21, 2020"
output: 
  html_document:
    keep_md: true
---

Recently, I have found a use tsCV from the forecast package.  However, when using the `xreg` option I encountered a few issues and decided to look into them.  This report summarizes the findings and proposed fixes for this function.  First, a great thank you to Rob J Hyndman and George Athanasopoulos for providing such a wonderful resource that is freely available.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(forecast)
```

For the practice data set, I have loaded a time series object and xreg matrix.  For more details on the subject matter, please refer to my [EIA Data Analysis Project](https://github.com/john-grando/eia-data-analysis).  For these purposes, it is only necessary to note that two predictor variables are provided in the `p_xreg` object.

If we try to use the tsCV function out of the box using xregs, we get an output of `NA` values

```{r}
my_file <- "https://github.com/john-grando/eia-data-analysis/blob/master/app/RFiles/Data/total_energy_coal_data.RData?raw=true"
load(url(my_file))
my_fun <- function(x, h){forecast(Arima(x, order=c(2,0,0)), h=h, xreg=xreg)}
(e <- tsCV(eng_coal_ts, my_fun, h=1, xreg=p_xreg))
```

To debug this problem, we can step through the function.  First, we will set the initial parameters

```{r}
y <- eng_coal_ts
h = 2
xreg = p_xreg
initial = 0
window = NULL
forecastfunction = my_fun
```

```{r}
y <- as.ts(y)
n <- length(y)
e <- ts(matrix(NA_real_, nrow = n, ncol = h))
if(initial >= n) stop("initial period too long")
tsp(e) <- tsp(y)
if (!is.null(xreg)) {
  # Make xreg a ts object to allow easy subsetting later
  xreg <- ts(as.matrix(xreg))
  if(NROW(xreg) != length(y))
    stop("xreg must be of the same size as y")
  tsp(xreg) <- tsp(y)
}
if (is.null(window)) {
  indx <- seq(1+initial, n - 1L)
} else {
  indx <- seq(window+initial, n - 1L, by = 1L)
}
```

Eventually we get to a for loop that contains the error.  Here we have set the looped value arbitrarily to 20.

```{r error=TRUE}
i = 20
y_subset <- subset(
    y,
    start = ifelse(is.null(window), 1L,
            ifelse(i - window >= 0L, i - window + 1L, stop("small window"))
    ),
    end = i
  )
  if (is.null(xreg)) {
    fc <- try(suppressWarnings(
      forecastfunction(y_subset, h = h, ...)
      ), silent = TRUE)
  } else {
    xreg_subset <- as.matrix(subset(
      xreg,
      start = ifelse(is.null(window), 1L,
              ifelse(i - window >= 0L, i - window + 1L, stop("small window")))
    ))
    # The `...` has been removed and xreg has been inserted into the arguments
    # for out-of-function operation.
    fc <- forecastfunction(y_subset, h = h, xreg = xreg_subset)
  }
```
The error is not immediately obvious.  If we want to use `xreg` in the training, then we also need to specify an `xreg` in the prediction forecast which is the same length as `h`.  Therefore, we need to identify a new parameter to identify the prediction `xreg` argument.

As we debug this operation, there are a few other issues that arise:
1. The `subset()` function that creates `xreg_subset` does not have an end, thus returning a matrix with [i:length(xreg)].  This can be fixed by  adding an `end=i` term. 
2.  The subset does not behave well with multidimensional arrays because when they are subset, they are converted to 1 dimensional arrays.  It is much easier to slice the matrix and reformat it to the original 2D form.

3.  If we want to get maximum results for each prediction of `h` for each index location, a different function should be fit for each `h`.

```{r}
#interval from y_subset
#start = ifelse(is.null(window), 1L,
#        ifelse(i - window >= 0L, i - window + 1L, stop("small window")))
#prediction_xreg = xreg[start:h,]

prediction_xreg <- as.matrix(subset(
      xreg,
      start = ifelse(is.null(window), 1L,
              ifelse(i - window >= 0L, i - window + 1L, stop("small window"))),
      end = h
    ))

    xreg_subset <- as.matrix(subset(
      xreg,
      start = ifelse(is.null(window), 1L,
              ifelse(i - window >= 0L, i - window + 1L, stop("small window"))),
      end = i
    ))

my_fun <- function(x, h, px, xr){forecast(Arima(x, order=c(2,0,0), xreg=xr), h=h, xreg=px)}
forecastfunction <- my_fun
fc <- forecastfunction(x = y_subset, h = h, px=prediction_xreg, xr=xreg_subset)
if (!is.element("try-error", class(fc))) {
    e[i, ] <- y[i + (1:h)] - fc$mean
}
print(e[20,])
```
The changes above, as well as a few formatting fixes (which may be due to an updated R version), have been applied to the function below

```{r}
tsCV_v2 <- function(y, forecastfunction, h=1, window=NULL, xreg=NULL, initial=0, console_print=NULL, ...) {
  y <- as.ts(y)
  n <- length(y)
  e <- ts(matrix(NA_real_, nrow = n, ncol = h))
  if(initial >= n) stop("initial period too long")
  tsp(e) <- tsp(y)
  if (!is.null(xreg)) {
    xreg <- as.matrix(xreg)
    if(NROW(xreg) != length(y))
      stop("xreg must be of the same size as y")
    tsp(xreg) <- tsp(y)
  }
  if (is.null(window)) {
    indx <- seq(1+initial, n - 1L)
  } else {
    indx <- seq(window+initial, n - 1L, by = 1L)
  }
 for (i in indx) {
    y_subset <- subset(
      y,
      start = ifelse(is.null(window), 1L,
              ifelse(i - window >= 0L, i - window + 1L, stop("small window"))
      ),
      end = i
    )
    if (is.null(xreg)) {
      fc <- try(suppressWarnings(
        forecastfunction(y_subset, h = h, ...)
        ), silent = TRUE)
    f_mean <- if (!is.element("try-error", class(fc))){fc$mean}
    max_h <- h
    } else {
      start = ifelse(is.null(window), 1L,
              ifelse(i - window >= 0L, i - window + 1L, stop("small window")))
      end = i
      xreg_subset <- matrix(xreg[start:end,], ncol = ncol(xreg))
      #process each h separately due to errors that can happen near the end of the index
      make_prediction_m <- function(h, i) {try(matrix(xreg[i:(i+h-1),], ncol = ncol(xreg)), silent = TRUE)}
      prediction_l <- lapply(1:h, make_prediction_m, i=i)
      fc_fun <- function(px, y_subset, xr, ...) {
          try(suppressWarnings(
            forecastfunction(y_subset, h = nrow(px), xr = xreg_subset, px = px, ...)
          ), silent = TRUE)
      }
      fc_l <- lapply(prediction_l, fc_fun, y_subset=y_subset, xr=xreg_subset)
      f_mean_l <- lapply(fc_l, function(x){if (!is.element("try-error", class(x))){x$mean}})
      max_h <- which.max(lapply(f_mean_l, length))
      f_mean <- f_mean_l[[max_h]]
    }
    if (!is.null(f_mean)) {
      tmp_result <- as.vector(y[i + (1:max_h)] - f_mean)
      result <- c(
        tmp_result, 
        rep(NA,h-length(tmp_result)))
      e[i, ] <- result
    }
  }
  if (h == 1) {
    return(e[, 1L])
  } else {
    colnames(e) <- paste("h=", 1:h, sep = "")
    return(e)
  }
}
```

Additionally, the new tsCV function has been vectorized, which removes all uses of for loops.  This should provide better performance for larger data sets.

```{r}
tsCV_v2_vectorized <- function(y, forecastfunction, h=1, window=NULL, xreg=NULL, initial=0, ...) {
  y <- as.ts(y)
  n <- length(y)
  e <- ts(matrix(NA_real_, nrow = n, ncol = h))
  if(initial >= n) stop("initial period too long")
  tsp(e) <- tsp(y)
  if (!is.null(xreg)) {
    xreg <- as.matrix(xreg)
    if(NROW(xreg) != length(y))
      stop("xreg must be of the same size as y")
    tsp(xreg) <- tsp(y)
  }
  if (is.null(window)) {
    indx <- seq(1+initial, n - 1L)
  } else {
    indx <- seq(window+initial, n - 1L, by = 1L)
  }
  y_subset_fun <- function(i){
    subset(
      y,
      start = ifelse(is.null(window), 1L,
              ifelse(i - window >= 0L, i - window + 1L, stop("small window"))
      ),
      end = i
    )
  }
  y_subset_l <- lapply(indx, y_subset_fun)
  if (is.null(xreg)) {
    fc_fun <- function(y_subset, h=h){try(suppressWarnings(
        forecastfunction(y_subset, h = h)
        ), silent = TRUE)}
    fc_l <- lapply(y_subset_l, fc_fun, h=h)
    f_mean_l <- lapply(fc_l, function(x){if (!is.element("try-error", class(x))){x$mean}})
    max_h <- h
  } else {
    xreg_subset_fun <- function(i){
      x_start = ifelse(is.null(window), 1L,
              ifelse(i - window >= 0L, i - window + 1L, stop("small window")))
      x_end = i
      matrix(xreg[x_start:x_end,], ncol = ncol(xreg))
    }
    #subset xreg for each indx
    xreg_subset_l <- lapply(indx, xreg_subset_fun)
    #subset xreg predictor for each h per each indx
    prediction_per_h_subset <- function(i, h){
        lapply(
          1:h, 
          function(x, i){
            try(matrix(xreg[i:(i+x-1),], ncol = ncol(xreg)), silent = TRUE)
          }, 
          i=i)
    }
    #make list of xreg predictors per indx
    prediction_l <- lapply(indx, prediction_per_h_subset, h=h)
    #run function and return list of xreg predictors per indx
    fc_fun <- function(i, y_subset_l, prediction_l, xreg_subset_l, ...) {
      prediction <- prediction_l[[i]]
      y_subset <- y_subset_l[[i]]
      xr = xreg_subset_l[[i]]
      lapply(
        prediction[1:h],
        function(px){
          try(suppressWarnings(
            forecastfunction(px = px, x = y_subset, h = nrow(px), xr = xr, ...)
          ), silent = TRUE)
        })
    }
    #From this point forward, process the lists using their order, which is different
    #than the raw indx due to window and initial inputs
    fc_l <- lapply(1:length(indx), fc_fun, y_subset_l=y_subset_l, prediction_l=prediction_l, xreg_subset_l=xreg_subset_l)
    #extract means from each predictor subset of indx
    f_mean_long_l <- lapply(
        fc_l, 
        function(x){
          lapply(
            1:h,
            function(x2){
              if (!is.element("try-error", class(x[[x2]]))){x[[x2]]$mean}}
          )
        })
    #find longest arrary of each subset and use that one for the indx
    max_l <- lapply(f_mean_long_l, function(x){which.max(lapply(x, length))})
    f_mean_l <- lapply(1:length(indx), function(x){f_mean_long_l[[x]][[max_l[[x]][[1]]]]})
    max_h <- h
  }
  #process errors
  errors_fun <- function(i){
    f_tmp <- c(
      f_mean_l[[i]],
      rep(NA,h-length(f_mean_l[[i]]))
    )
    tmp_result <- as.vector(y[i + (1:max_h)] - f_tmp)
    result <- c(
      tmp_result, 
      rep(NA,h-length(tmp_result)))
  }
  errors_l <- lapply(1:length(indx), errors_fun)
  window <- ifelse(is.null(window),0,window)
  lapply(1:length(indx), function(x){e[x+window+initial,] <<- errors_l[[x]]})
  if (h == 1) {
    return(e[, 1L])
  } else {
    colnames(e) <- paste("h=", 1:h, sep = "")
    return(e)
  }
}
```

Test using no `xreg`.  This allows the original function to be compared to the two new options

```{r}
my_fun <- function(x, h){forecast(Arima(x, order=c(2,0,0)), h=h)}

e1_start <- proc.time()
e1 <- tsCV(eng_coal_ts, my_fun, h=3)
original_function_duration <- proc.time() - e1_start
e2_start <- proc.time()
e2 <- tsCV_v2(eng_coal_ts, my_fun, h=3)
fixed_duration <- proc.time() - e2_start
e3_start <- proc.time()
e3 <- tsCV_v2_vectorized(eng_coal_ts, my_fun, h=3)
fixed_vector_duration <- proc.time() - e3_start
all.equal(e1,e2, e2, tolerance=0.1)
original_function_duration
fixed_duration
fixed_vector_duration
```

Test using `xreg`.  Note, the original function could not be compared, as it provides erroneous output

```{r}
my_fun <- function(px, x, h, xr){forecast(Arima(x, order=c(2,0,0), xreg=xr), h=h, xreg=px)}
e2xr_start <- proc.time()
e2xr <- tsCV_v2(eng_coal_ts, my_fun, h=3, xreg=p_xreg)
fixed_duration <- proc.time() - e2xr_start
e3xr_start <- proc.time()
e3xr <- tsCV_v2_vectorized(eng_coal_ts, my_fun, h=3, xreg=p_xreg)
fixed_vector_duration <- proc.time() - e3xr_start
all.equal(e2xr, e2xr, tolerance=0.1)
fixed_duration
fixed_vector_duration
```

